---
title: "Class 10 Non Parametric Test and Regression Analysis"
author: "Md Rana Mahmud"
date: "9/19/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

The one-sample Wilcoxon signed rank test is a non-parametric alternative to one-sample t-test when the data cannot be assumed to be normally distributed. It’s used to determine whether the median of the sample is equal to a known standard value (i.e. theoretical value).

```{r}
set.seed(1234)
df <- data.frame(
  weight = round(rnorm(10, 25, 5), 1)
)
df
# We want to know, if the average weight of the mice differs from 25g (two-tailed test)?

# 5% alpha
wilcox.test(df$weight, mu = 25)
```


Wilcoxon Rank Sum test' (also called 'Mann-Whitney test'), is a distribution-free alternative to the t-test, and is used to test the hypothesis that the distributions in the two groups have the same median.

```{r}
## Two-sample test.
## Hollander & Wolfe (1973), 69f.
## Permeability constants of the human chorioamnion (a placental
##  membrane) at term (x) and between 12 to 26 weeks gestational
##  age (y).  The alternative of interest is greater permeability
##  of the human chorioamnion for the term pregnancy.
x <- c(0.80, 0.83, 1.89, 1.04, 1.45, 1.38, 1.91, 1.64, 0.73, 1.46)
y <- c(1.15, 0.88, 0.90, 0.74, 1.21)
wilcox.test(x, y, alternative = "g")        # greater
wilcox.test(x, y, alternative = "greater",
            exact = FALSE, correct = FALSE) # H&W large sample
                                            # approximation

```
```{r}
## Hollander & Wolfe (1973), 29f.
## Hamilton depression scale factor measurements in 9 patients with
##  mixed anxiety and depression, taken at the first (x) and second
##  (y) visit after initiation of a therapy (administration of a
##  tranquilizer).
x <- c(1.83,  0.50,  1.62,  2.48, 1.68, 1.88, 1.55, 3.06, 1.30)
y <- c(0.878, 0.647, 0.598, 2.05, 1.06, 1.29, 1.06, 3.14, 1.29)
wilcox.test(x, y, paired = TRUE, alternative = "greater")
wilcox.test(y - x, alternative = "less")    # The same.
wilcox.test(y - x, alternative = "less",
            exact = FALSE, correct = FALSE) # H&W large sample
                                            # approximation
```


Regression Analysis
1. linear regression (a) simple, (b) multiple
2. logistic regression (a) binary , (b) multiple/multinomial logistic regression
y contionus linear regression
y categorical x contious/ categorical logistic regression
y <- x1 + x2 
```{r}
# lm()

data(iris)
y ~ x1 + x2 + x3
reg_model <- lm(Sepal.Width ~ Sepal.Length, data = iris)

summary(reg_model)
```
```{r}
par(mfrow=c(2,2))
plot(reg_model)
```


1. Multicollinearity 
2. Residuals are normally distributed
3. Homogenity of varaince
4. High influential obsrevation
VIF value that exceeds 5 or 10 indicates a problematic amount of collinearity (James et al. 2014).
```{r}
library(car)
fit <- lm(prestige ~ income + education + type, data=Duncan)
vif(fit)
```

