---
title: "Day 9 Hypothesis Testing"
author: "Md Rana Mahmud"
date: "9/18/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```
# Inference concerning the mean

## Normal population-varaince known

Suppose a sample of size n from normal population with mean $\mu$ and known variance $\sigma^2$

$H_{0}: \mu = \mu_{0}$
$H_{1}: \mu \ne \mu_{0}$

if population not normal we can use this test if n is large >30

Shapiro-Wilk normality test


average longevity of bulb is 4 years, standard deviation 0.16, random sample of 40 gave mean 3.45
Two tailed test
example mn islam 11.5
alpha = 0.05
```{r}
# normal test
x_bar <- 3.45
n <- 40
mu_not <- 4
sigma <- 0.16
z_statistics  <- (x_bar - mu_not)/(sigma/sqrt(n))
z_statistics
# 0.05
# 0.025
# 1- 0.025
qnorm(c(0.025, 0.975))
```
```{r}
# critical value -1.96
2*pnorm(-abs(z_statistics))

```
Reject null hypothesis

```{r}
mu_not - 1.96*(sigma/sqrt(n))
mu_not + 1.96*(sigma/sqrt(n))
```
1. compairing test statistics
2. confidence interval
3. p value calculate

## normal population variance unknown-sample size small


$t = \dfrac{\overline{x} - \mu_{0}}{s/\sqrt(n)}$
with n-1 df
t distribution instead of normal

example mn inslam book 11.11

tensile strength of steel averge 120 pounds, sample 25, sample mean 110 pounds, sample variance 144

```{r}
x_bar <- 110
n <- 25
mu_not <- 120
s <- 12
t_statistics  <- (x_bar - mu_not)/(s/sqrt(n))
t_statistics

# from t distribution
# p value
2*pt(-abs(t_statistics),df=n-1)
# critical region
qt(c(.025, .975), df=n-1)


```
reject null hypothesis

## difference between two mean

### testing equality of two population means

assumptions populations are normal with known but unequal variance
drawn independentlly from the populations

example 11.14

trival women 12, non trival 15, average children respectively 4.5, 3.4,population variance 1 and 1.5 respectively
```{r}
n1 <- 12
n2 <- 15
x1_bar <- 4.5
x2_bar <- 3.4
sigma_sq_1 <- 1
sigma_sq_2 <- 1.5

z <- (x1_bar-x2_bar)/(sqrt((sigma_sq_1/n1)+sigma_sq_2/n2))
z
# critical region
qnorm(c(0.025,0.975))


2*pnorm(-abs(z))
```

Population Variance unknown small sample
assuming two indepndent population

example 11.18

batsman average 70.2, sample standard deviation 8.4, in 16 matches in dhaka, 62.7 std 7.7 in another 14 match in chittagong

difference 7.5 due to regional difernce or real difference

```{r}
n1 <- 16
n2 <- 14
x1_bar <- 70.2
x2_bar <- 62.7
sigma_sq_1 <- 8.4^2
sigma_sq_2 <- 7.5^2

sp_sq <- ((n1-1)*sigma_sq_1 + (n2-1)*sigma_sq_2)/(n1+n2-2)
sp_sq


t <- (x1_bar -x2_bar)/(sqrt(sp_sq)*sqrt(1/n1+1/n2))
t
# critical region
qt(c(0.025,0.975), n1+n2-2)


pt(-abs(t),df=n1+n2-2)

```
reject null hypothesis

bating average of two places is due to chance is less than  5 percent.

# paired comparison

example 11.20

```{r}
sales99 <- c(126, 56, 86, 62, 96, 36, 52, 50, 35, 53)
sales20 <- c(123, 49, 79, 59, 92, 35, 48, 48, 32, 50)

d1 <- sales99 - sales20
d1
d1_sq <- d1^2
d_bar <- sum(d1)/length(sales99)
n <- length(sales99)
d_bar
s_d <- sqrt((sum(d1^2) - n * d_bar^2)/(n-1))
s_d

t <- d_bar / (s_d / sqrt(n))
t

qt(c(0.025,0.975), n-1)

```

reject null hypothesis

using r command for independent t test
```{r}
t.test(sales99, sales20, paired = TRUE)
# less than
t.test(sales99, sales20, paired = TRUE, alternative = "less")
t.test(sales99, sales20, paired = TRUE, alternative = "greater")
# changing alpah value
t.test(sales99, sales20, paired = TRUE, conf.level = 0.99)
t.test(sales99, sales20, paired = TRUE, conf.level = 0.90)
```

normality test
shapiro wilk test
 null-hypothesis of this test is that the population is normally distributed.
```{r}
shapiro.test(sleep[sleep$group == 1, "extra"])
shapiro.test(sleep[sleep$group == 2, "extra"])
```

 Independent t test in R
```{r}
data("sleep")
t.test(extra ~ group, data = sleep)
x1 <- c(1,2, 7, 9,8,7,29,23)
x2 <- c(8,6, 99,2,2,7,19,8)
df <- data.frame(x1, x2)
t.test(x1, x2)
t.test(x1, x2, data = df)
```
 
## Test concerning proportions

$H_{0}: p = p_{0}$
$H_{1}: p => p_{0}$

example 11.28

protien deficiency at most 0.75, true for 244 out of 300. sig 0.01
```{r}
p_caret <- round(244/300,2)
p_not <- 0.75
q_not <- 1 - p_not
n <- 300
z <- (p_caret - p_not)/sqrt(p_not*q_not/n)
z

abs(qnorm(0.01))
```

```{r}
heads <- rbinom(1, size = 100, prob = .5)
prop.test(heads, 100)          # continuity correction TRUE by default
prop.test(heads, 100, correct = FALSE)

## Data from Fleiss (1981), p. 139.
## H0: The null hypothesis is that the four populations from which
##     the patients were drawn have the same true proportion of smokers.
## A:  The alternative is that this proportion is different in at
##     least one of the populations.

smokers  <- c( 83, 90, 129, 70 )
patients <- c( 86, 93, 136, 82 )
prop.test(smokers, patients)

prop.test(244/300,p = 0.75, n = 300)

```

compairing single mean

```{r}
mean(sleep$extra)

t.test(sleep$extra, mu = 1.5)
```

# chisquare test

```{r}
data("mtcars")
```
null hypothesis : cylindar and am are independent

```{r}
chisq.test(mtcars$cyl, mtcars$am)
chisq.test(mtcars$cyl, mtcars$am)
```

```{r}
table(mtcars$cyl, mtcars$am)
```

# fishers exact test

```{r}
## Agresti (1990, p. 61f; 2002, p. 91) Fisher's Tea Drinker
## A British woman claimed to be able to distinguish whether milk or
##  tea was added to the cup first.  To test, she was given 8 cups of
##  tea, in four of which milk was added first.  The null hypothesis
##  is that there is no association between the true order of pouring
##  and the woman's guess, the alternative that there is a positive
##  association (that the odds ratio is greater than 1).


TeaTasting <-
matrix(c(3, 1, 1, 3),
       nrow = 2,
       dimnames = list(Guess = c("Milk", "Tea"),
                       Truth = c("Milk", "Tea")))
chisq.test(TeaTasting)
chisq.test(TeaTasting,simulate.p.value = TRUE)
fisher.test(TeaTasting, alternative = "greater")
## => p = 0.2429, association could not be established
```

# Data Subset

```{r}
library(dplyr)
```

```{r}
df2 <- mtcars %>%
    # variable selection
    select(vs, am)
head(df2)
df3 <- mtcars %>%
    # variable selection
    select(-vs, -am)
head(df3)
# starts_with
df4 <- mtcars %>%
    # variable selection
    select(starts_with("d"))
head(df4)

```

```{r}
# filter
df5 <- mtcars %>%
    filter(mpg >= 25)
df5

mtcars %>%
    filter(mpg >= 25) %>%
    select(mpg, cyl)
```

```{r}
# summarize
mtcars %>% 
    group_by(cyl) %>% 
    summarise(avg_mpg =  mean(mpg), sd_mpg = mean(mpg))
mtcars %>% 
    group_by(cyl,am) %>% 
    summarise(avg_mpg =  mean(mpg), sd_mpg = mean(mpg))
df6 <- mtcars %>% 
    filter(mpg > 25)%>%
    group_by(cyl,am) %>% 
    summarise(avg_mpg =  mean(mpg), sd_mpg = mean(mpg))
df6
```

```{r}
data(airquality)
subset(airquality, Temp > 80)
subset(airquality, Temp > 80, select = c(Ozone, Temp))
subset(airquality, Temp > 80, select = -Temp)
subset(airquality, Temp > 80, select = Ozone: Wind)
```


```{r}
library(ggplot2)
# base ggplot()
# aesthetics / line type
# theme
# geom_line()
# geom_hist()
# geom_bar()
# stat_
data("diamonds")
# aes(x = , y = , color, size, alpha )
# scatterplot = geom_point
ggplot(diamonds) + geom_point( aes(depth, price))
ggplot(diamonds) + geom_point( aes(depth, price, alpha = 0.25))+ggtitle("Relationship Between Depth and Price")+xlab("Depth")+ylab("Price")
```


